Citation Analysis of Computer Science

### 2019-11-17 Scientometrics/comp.tex
wrote up results for Fig 1

In extending the work of Salton and Bergmark to the 21st century, we sought to combine journal and article based classifications in a scalable manner. Consequently, we used a whole text search method to match records from two major data sources. (i) the dblp bibliography- a reference bibliography for computer science~\cite{dblp_ref} (ii) the journal based Scopus bibliography, which more comprehensively covers the physical sciences, life sciences, social sciences, and health sciences. We assumed that a record in dblp qualifies for a computer science label and cross-matched dblp publications to Scopus to  enable identification of citing and cited publications across multiple disciplines.

@@

\emph{Data} A stable release of the dblp computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the dblp dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning vectors, containing these lexemes. The vectors are then matched with corresponding text search vectors extracted from DBLP document titles. A rank is returned based on the frequency of matching lexemes. Only recordxs in dblp with DOIs (4,291,130) were considered and only matches with a normalized rank of $>$ 0.99 were accepted, roughly equivalent to `at least 11 matching lexemes independent of order and vector length'. As an example, the following query returns a normalized rank of 0.994.
\begin{lstlisting}
SELECT ts_rank(to_tsvector(`english', `one two three 
four five six seven eight nine ten eleven'), 
plainto_tsquery(`english', `one four three two five 
six seven eight nine ten eleven')); 
\end{lstlisting}

 This procedure resulted in a dataset of 634,100 dblp publications with Scopus identifiers. Cited references for these were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations of these references and publications within the dataset. This dataset is referred to as \emph{comp}.\par

\emph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}. We downloaded v1.2 from the authors' website at `cs.utexas.edu'. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus,  The number of clusters to be formed is required by Graclus, and we varied this parameter between 10 and 50 clusters. At a setting of 20, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. [@sitaram please add the number of nodes in the Graclus 20 clusters].

@@

For the purpose of this study, we defined, the computer science literature as a working set of all articles in the dblp bibliography published in 1996-2015, that had a digital object identifier (DOI), and could be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 634,100 publications (Materials and Methods). We constructed article level clusters of the computer science literature for the years 1996-2015, at a sufficiently high level of aggregation to support cognitive ease. Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using a spectral clustering technique, Graclus, that has previously been applied to citation data~\cite{graclus_2007, subelj_clustering_2016}.  \emph{@sitaram please confirm the number of pubs that went into Graclus. I'm assuming it was 634,100?}. To place these clusters in context, we identified that accounted for at least 10\% of the publications in each cluster and observed 31 of the 334 minor subject areas in the Scopus ASJC classification. (which consists of 5 subject areas, 27 major subject areas, and 334 minor subject areas).

>>>>

In extending the work of Salton and Bergmark to the 21st century, we sought to combine journal and article based classifications in a scalable manner. Consequently, we used a whole text search method to match records from two major data sources. (i) the dblp bibliography- a reference bibliography for computer science~\cite{dblp_ref} (ii) the journal based Scopus bibliography, which more comprehensively covers the physical sciences, life sciences, social sciences, and health sciences. The underlying assumption is that a record in dblp qualifies as computer science. Therefore, we cross-matched dblp publications to Scopus to  enable identification of linked publications across multiple disciplines.

@@

\emph{Data} A stable release of the dblp computer science bibliography~\cite{dblp_ref} consisting of 7,079,994 records was downloaded as dblp-2018-08-01.xml.gz. Publications were parsed from the xml source file and loaded into a PostgreSQL database. We have previously parsed, the Scopus dataset of over 72,000,000 publications into a custom schema as part of a larger data platform for research evaluation~\cite{GithubERNIE2019}. Records in the dblp dataset, were matched to Scopus identifiers using PostgreSQL's full text search feature applied to document titles. The process involves parsing all Scopus document titles into tokens, reducing the tokens into lexemes and returning vectors containing these lexemes. The vectors are then matched with corresponding text search vectors extracted from DBLP document titles. A rank is returned based on the frequency of matching lexemes. Only recordxs in dblp with DOIs (4,291,130) were considered and only matches with a normalized rank of $>$ 0.99 were accepted, roughly equivalent to `at least 11 matching lexemes independent of order and vector length'. As an example, the following query returns a normalized rank of 0.994.
\begin{lstlisting}
SELECT ts_rank(to_tsvector(`english', `one two three 
four five six seven eight nine ten eleven'), 
plainto_tsquery(`english', `one four three two five 
six seven eight nine ten eleven')); 
\end{lstlisting}

This procedure resulted in a dataset of 634,100 dblp publications with Scopus identifiers. Cited references were extracted from Scopus, resulting in a total of 3,960,503 nodes (publications and references) and 11,992,387 edges that represent citations of these references and publications within the dataset. This dataset is referred to as \emph{comp}.\par

\emph{Clustering by Direct Citation} Graclus is a spectral graph clustering package that computes normalized cut and ratio association for a given undirected graph~\cite{graclus_2007}. We downloaded v1.2 from the authors' website at `cs.utexas.edu'. The \emph{comp} dataset was formatted as an undirected graph, stored in a file with a header line indicating the number of nodes and edges, and used as input to Graclus,  We varied the number of clusters to be formed between 10 and 50 clusters. At 20 clusters, the most evenly sized set of clusters resulted with the largest cluster containing roughly 10 times the number of nodes in the smallest one~\cite{traag_louvain_2019}. [@sitaram please add the number of nodes in the Graclus 20 clusters].

@@

For the purpose of this study, we defined, the computer science literature as a working set of all articles in the dblp bibliography published in 1996-2015, that had a digital object identifier (DOI), and could be matched to article identifiers in the Scopus bibliography. The process of selection and matching resulted in a dataset of 634,100 publications (Materials and Methods). We constructed article level clusters of the computer science literature for the years 1996-2015, at a sufficiently high level of aggregation to be comparable to disciplinary classfication by journal. Clustering of publications is commonly accomplished through direct citation, bibliographic coupling, and co-citation with an argument being made for direct citation as the best approach to concentrate citation links~\cite{klavans_which_2017}. Accordingly, we first used direct citation links as the basis for cluster formation. A set of 20 clusters was generated using a spectral clustering technique, Graclus, that has previously been applied to citation data~\cite{graclus_2007, subelj_clustering_2016}.  \emph{@sitaram please confirm the number of pubs that went into Graclus. I'm assuming it was 634,100?}. To place these clusters in context, we identified that accounted for at least 10\% of the publications in each cluster and observed 31 of the 334 minor subject areas in the Scopus ASJC classification. (which consists of 5 subject areas, 27 major subject areas, and 334 minor subject areas).
